---
title: 'EDA Project: Diamond Prices (Part 1)'
author: "Sean Anthony"
date: "February 17, 2016"
output: html_document
---

```{r}
install.packages('ggplot2')
library(ggplot2)
data(diamonds)
summary(diamonds)

setwd("/Users/sa1251/Documents/eda-project/eda-project")
getwd()
```

# Create a histogram of the price of
# all the diamonds in the diamond data set.

```{r}
qplot(x = price, data = diamonds)
ggsave('price_diamonds.png')
```

# Explore the largest peak in the
# price histogram you created earlier.

```{r}
qplot(x = price, data = diamonds, binwidth = 100, color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 100))
ggsave('largest_peak.png')
```

# Break out the histogram of diamond prices by cut.

```{r}
qplot( x = price, data = diamonds, color = I('black'), fill = I('#F79420')) + 
  facet_wrap(~cut)
ggsave('prices_by_cut.png')
```

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales = "free_y")
ggsave('facet_free_y.png')
```

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

```{r}
qplot(x = price/carat, data = diamonds, binwidth = .05, color = I('black'), fill = I('#F79420')) + 
  facet_wrap(~cut) + scale_x_log10(limits = c(1000, 12000), breaks = seq(0, 10000, 2000))
ggsave('price_per_carat.png')
```

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

```{r}
by(diamonds$price, diamonds$color, summary)
qplot(x = color, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim = c(1000, 8000))
ggsave('diamonds_boxplot.png')

```

# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

```{r}
qplot(x = color, y = price/carat, data = diamonds, geom = 'boxplot') + 
  scale_y_log10(breaks = seq(2000, 10000, 1000))
ggsave('price_per_carat_boxplot.png')
```

# Gapminder Data

```{r}

library(ggplot2)
library(dplyr)
library(tidyr)

life_exp <- read.csv('life_exp.csv', header = T, check.names = F)
life_exp <- gather(life_exp, "year", "n", 2:217)


# This graph shows the distribution of all countries' average life expectancies over the last 
# decade.

ggplot(aes(x = n, binwidth = 1), data = subset(life_exp, year >= 2006)) + 
  geom_histogram(color = I('black'), fill = I('#F79420')) + xlab('Average Life Expectancy') + 
  scale_x_continuous(limits = c(40, 90), breaks = seq(40, 90, 2))
ggsave('last_decade_histogram.png')

# I've highlighted this same data using different tools. This dotplot faceted by every year may
# do a better job highlighting yearly differences. 

ggplot(aes(x = n), data = subset(life_exp, year >= 2006)) + geom_dotplot() + 
  xlab('Average Life Expectancy') + facet_wrap(~year, ncol = 5) + 
  scale_x_continuous(limits = c(40,90), breaks = seq(40, 90, 5))
ggsave('last_decade_dotplot.png')

```

# Facebook Birthday Data

```{r}

# Using the sample data set

bday_data <- read.csv('birthdaysExample-2.csv')
library(dplyr)
library(tidyr)
bday_data <- as.Date(bday_data$dates, format = "%m/%d/%y")
bday_data <- as.data.frame(bday_data)
colnames(bday_data)[1] <- 'dates'
bday_data <- separate(bday_data, dates, c("y", "m", "d"))
colnames(bday_data)[1:3] <- c('year', 'month', 'day')

# How many people share your birthday?
filter(bday_data, month == 12, day == 10)

# Which month contains the most number of birthdays?
table(bday_data$month)
max(table(bday_data$month))

# How many birthdays are in each month?
table(bday_data$month)

# Which day of the year has the most number of birthdays?
table(bday_data$day)
max(table(bday_data$day))

# Do you have at least 365 friends that have birthdays on every day of the year?
distinct(bday_data)
```

